---
title: "Rohan Bandaru - Ch1 Final Submission GOAT Project"
output: html_notebook
---

This is an R notebook. It is my first time using R and I learned a lot about
practical application of statistics. I also used Python for some 
pre-processing (filtering csv files) allowing me to use hundreds of data points.

# Intro

I chose tennis as my sport and listed the following players as contenders for
the "Greates Of All Time" title. 

Roger Federer	 -SUI
Novak Djokovic -SRB
Rafael Nadal	 -ESP
Ivan Lendl	   -USA
Jimmy Connors	 -USA
Pete Sampras	 -USA
John McEnroe	 -USA
Bjorn Borg	   -SWE
Andre Agassi	 -USA
Borris Becker	 -GER

There are a few things I wanted to consider when choosing the GOAT:

- game skill
- how long they were good
- consistency (inter-game variability)

These qualities are what led me to choose the statistics below:

No. Grand Slams Won, No. Big Titles Won (GS, TF, M), Games Played, Games Won, 
Matches Played, Matches Won, Weeks at #1, Hard Court Win %, Clay Court Win %, 
Grass Court Win %, Carpet Court Win %, Elo (rating), Match time
*I didn't use all these statistics in my final analysis because there was no
easy way I could pull come of them off the website

The reason why I am not doing win rate or win/loss ratio, is because I believe
that the title of GOAT should be based on the total amount of wins and titles 
accumulated over time. A player may win all his matches but if he only plays in
one title, he should not become the GOAT. 

I also gathered the raw data (score, surface, and duration) from every game 
they ever played (that was recorded).

To compare the players, I will take a weighted combination of their Z-scores 
for each statistic

Source: https://www.ultimatetennisstatistics.com/

# Setting Up Environment

```{r}
setwd("~/Documents/AP Stats")
```

# Statistic 1 - Number of Grand Slam Titles Won

Grand Slam Titles are the highest level of competition in tennis: 
US Open, Australian Open, French Open(Roland Garros), and Wimbledon. 

```{r}
overallstats <- read.csv(file = 'overallstats.csv')

grandslams <- overallstats[["GrandSlams"]]
names(grandslams) <- overallstats[["PlayerName"]]

par(mar=c(7.5, 4.1, 1, 1))
barplot(sort(grandslams, decreasing = TRUE), las=3, ylab="GS Titles Won")
```
Calculate Z-Scores (Normalize Distribution)
```{r}
m = mean(grandslams)
stdv = sd(grandslams)

print(paste0("Mean GS Titles Won: ", m))
print(paste0("Standard Deviation: ", stdv))
print("Z-scores:")
# zs = (grandslams - m)/stdv
# summary(zs)
zscores_gs = scale(grandslams)
print(zscores_gs)
```
Total GS Titles Won In The Entire Group
```{r}
stem(grandslams)
```
A lot of the players have below 10 titles, and it quickly narrows down to the 
few top-players. 

# Statistic 2 - Number of Big Titles Won

Big Titles include Grand Slams(Majors), Tour Finals, Olympic Titles, etc. 

```{r}
bigtitles <- overallstats[["BigTitles"]]
names(bigtitles) <- overallstats[["PlayerName"]]

par(mar=c(7, 4.1, 1, 1))
barplot(sort(bigtitles, decreasing = TRUE), las=3, ylab="Big Titles Won")
```
Calculate Z-Scores (Normalize Distribution)
```{r}
m = mean(bigtitles)
stdv = sd(bigtitles)

print(paste0("Mean Big Titles Won: ", m))
print(paste0("Standard Deviation: ", stdv))
print("Z-scores:")
zscores_bt = scale(bigtitles)
print(zscores_bt)
```

# Statistic 3 - Games Won

```{r}
gameswon <- overallstats[["GamesWon"]]
names(gameswon) <- overallstats[["PlayerName"]]

par(mar=c(7, 4.1, 1, 1))
barplot(sort(gameswon, decreasing = TRUE), las=3, ylab="Games Won")
```
Calculate Z-Scores (Normalize Distribution)
```{r}
m = mean(gameswon)
stdv = sd(gameswon)

print(paste0("Mean Games Won: ", m))
print(paste0("Standard Deviation: ", stdv))
print("Z-scores:")
zscores_gw = scale(gameswon)
print(zscores_gw)
```

# Statistic 4 - Matches Won

```{r}
matcheswon <- overallstats[["MatchesWon"]]
names(matcheswon) <- overallstats[["PlayerName"]]

par(mar=c(7, 4.1, 1, 1))
barplot(sort(matcheswon, decreasing = TRUE), las=3, ylab="Matches Won")
```
Calculate Z-Scores (Normalize Distribution)
```{r}
m = mean(matcheswon)
stdv = sd(matcheswon)

print(paste0("Mean Matches Won: ", m))
print(paste0("Standard Deviation: ", stdv))
print("Z-scores:")
zscores_mw = scale(matcheswon)
print(zscores_mw)
```

# Statistic 5 - Weeks At #1 Rank

```{r}
weeksatone <- overallstats[["WeeksAt1"]]
names(weeksatone) <- overallstats[["PlayerName"]]

par(mar=c(7, 4.1, 1, 1))
barplot(sort(weeksatone, decreasing = TRUE), las=3, ylab="Weeks At #1")
```
Calculate Z-Scores (Normalize Distribution)
```{r}
m = mean(weeksatone)
stdv = sd(weeksatone)

print(paste0("Mean Weeks At The #1 Rank: ", m))
print(paste0("Standard Deviation: ", stdv))
print("Z-scores:")
zscores_wao = scale(weeksatone)
print(zscores_wao)
```

# Statistic 6 - Average Game Score Difference (Variability)

## Roger Federer
```{r}
RFmatches <- read.csv(file = 'Tennis Player Matches/1RFMatches.csv')
gamedifferencepermatch <- RFmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Federer -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Novak Djokovic
```{r}
NDmatches <- read.csv(file = 'Tennis Player Matches/2NDMatches.csv')
gamedifferencepermatch <- NDmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Djokovic -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Rafael Nadal
```{r}
RNmatches <- read.csv(file = 'Tennis Player Matches/3RNMatches.csv')
gamedifferencepermatch <- RNmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Nadal -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Ivan Lendl
```{r}
ILmatches <- read.csv(file = 'Tennis Player Matches/4ILMatches.csv')
gamedifferencepermatch <- ILmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Lendl -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Jimmy Connors
```{r}
JCmatches <- read.csv(file = 'Tennis Player Matches/5JCMatches.csv')
gamedifferencepermatch <- JCmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Connors -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Pete Sampras
```{r}
PSmatches <- read.csv(file = 'Tennis Player Matches/6PSMatches.csv')
gamedifferencepermatch <- PSmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Sampras -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## John McEnroe
```{r}
JMmatches <- read.csv(file = 'Tennis Player Matches/7JMMatches.csv')
gamedifferencepermatch <- JMmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "McEnroe -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Bjorn Borg
```{r}
BBmatches <- read.csv(file = 'Tennis Player Matches/8BBMatches.csv')
gamedifferencepermatch <- BBmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Borg -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Andre Agassi
```{r}
AAmatches <- read.csv(file = 'Tennis Player Matches/9AAMatches.csv')
gamedifferencepermatch <- AAmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Agassi -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```

## Borris Becker
```{r}
BKmatches <- read.csv(file = 'Tennis Player Matches/10BBckMatches.csv')
gamedifferencepermatch <- BKmatches[["GameDifference"]]
hist(gamedifferencepermatch, xlab = "Difference In Games Won", main = "Becker -- Games Won-Lost Every Match", breaks = 15)
```
Calculate "Variability" And Notice Distribution Shape
```{r}
m = mean(gamedifferencepermatch)
stdv = sd(gamedifferencepermatch)

summary(gamedifferencepermatch)
print(paste0("Standard Deviation: ", stdv))
```
Distribution Shape:

I noticed that the distributions seen above all had a very similar shape. 
One tall hump on the positive side of 0, and a smaller flat hump on the 
negative side. I think the reason for these two humps is because a player can 
either win or lose a match. If he won a match, he would most likely have won 
more games than his opponent and vice versa. Overall the distributions are very 
skewed left which is to be expected considering the players win/loss ratio. 
Because all the shapes are so similar, the only way I really compared them was 
by the height of the humps (# games won) in which case Federer is the winner. 

Normalize (Take Z-Scores)
```{r}
zs = c(4.80976431016085, 4.92418661319707, 5.11033557083605, 5.34604513444406, 5.28898042288614, 4.90347635797912, 5.1580742185318, 5.69334612535673, 5.662244099766, 5.20370473981154)
names(zs) = overallstats[["PlayerName"]]

scale(zs)

```


# Final Tallying

Weights I Determined:
Weight For Grand Slam Titles - 0.9
Weight For Big Titles - 0.8
Weight For Games Won - 0.7
Weight For Matches Won - 0.9
Weight For Weeks at #1 - 0.9
Weight For Game Winning Consistency(sd) - 0.5

Grand Slams, Big Titles, Games Won, Matches Won, Weeks at #1, Game Winning Consistency
Roger Federer	 -SUI
1.5508857 * 0.9 + 1.30636863 * 0.8 + 1.6581025 * 0.7 + 1.4562820 * 0.9 + 1.08405782 * 0.9 - 1.3330217 * 0.5
= 5.221358772
Novak Djokovic -SRB
0.9834885 * 0.9 + 1.46568187 * 0.8 - 0.1711528 * 0.7 - 0.1005984 * 0.9 + 0.80404288 * 0.9 - 0.9519427 * 0.5
= 2.095006868
Rafael Nadal	 -ESP
1.3617533 * 0.9 + 1.38602525 * 0.8 + 0.0559132 * 0.7 + 0.2682625 * 0.9 + 0.07400395 * 0.9 - 0.3319809 * 0.5
= 2.515586765
Ivan Lendl	   -USA
-0.7187031 * 0.9 - 0.04779397 * 0.8 + 0.6506236 * 0.7 + 0.6658658 * 0.9 + 0.68403648 * 0.9 + 0.4530405 * 0.5
= 1.211800856
Jimmy Connors	 -USA
-0.7187031 * 0.9 - 0.76470359 * 0.8 + 1.3992183 * 0.7 + 1.5472996 * 0.9 + 0.65403488 * 0.9 + 0.2629887 * 0.5
= 1.83355253
Pete Sampras	 -USA
0.4160913 * 0.9 - 0.44607709 * 0.8 - 0.6125576 * 0.7 - 0.8143682 * 0.9 + 0.84404502 * 0.9 - 1.0209174 * 0.5
= -0.894919384
John McEnroe	 -USA
-0.9078355 * 0.9 - 0.28676384 * 0.8 - 0.3640866 * 0.7 - 0.2443105 * 0.9 - 0.31601685 * 0.9 - 0.1729893 * 0.5
= -1.892112907
Bjorn Borg	   -SWE
-0.1513059 * 0.9 - 0.68504696 * 0.8 - 1.5908215 * 0.7 - 1.4275396 * 0.9 - 0.92604939 * 0.9 + 1.6097128 * 0.5
= -3.111161619
Andre Agassi	 -USA
-0.7187031 * 0.9 - 0.84436021 * 0.8 - 0.1430950 * 0.7 - 0.2970049 * 0.9 - 1.00605366 * 0.9 + 1.5061287 * 0.5
= -1.842175812
Borris Becker	 -GER
-1.0969679 * 0.9 - 1.08333008 * 0.8 - 0.8821442 * 0.7 - 1.0538883 * 0.9 - 1.89610113 * 0.9 - 0.0210186 * 0.5
= -5.136935901

# Visualization Of Ranking
```{r}
players = c(5.221358772, 2.095006868, 2.515586765, 1.211800856, 1.83355253, -0.894919384, -1.892112907, -3.111161619, -1.842175812, -5.136935901)
names(players) = overallstats[["PlayerName"]]
par(mar=c(7, 4.1, 1, 1))
barplot(sort(players, decreasing = TRUE), las=3, ylab="Rohan's Tennis GOAT Score")
```
So in the end, Roger Federer is officially the GOAT of Mens Tennis 
(according to me). I actually got surprisingly similar results to the 
website that I got my data from which also has a GOAT ranking. 
